The Effect of Events on Australian Parliamentary Discussion (1851--2017)

### 2018-10-04 outputs/docs/events/paper.Rmd
Minor updates to paper

---
output:
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: false
    fig_caption: true
    number_sections: true
    latex_engine: pdflatex
    template: svm-latex-ms.tex
title: "The Effect of Events on Australian Parliamentary Discussion (1851--2017)"
thanks: "Thank you to John Tang, Zach Ward, Tim Hatton, Martine Mariotti, Tianyi Wang, Matt Jacob, Leslie Root, Jill Sheppard, Matthew Kerby, and Chris Cochrane for their helpful suggestions. **Version as of**: `r format(Sys.time(), '%d %B %Y')`; **comments welcome**: rohan.alexander@anu.edu.au."
author:
- name: Monica Alexander
  affiliation: University of Toronto
- name: Rohan Alexander
  affiliation: Australian National University
abstract: "We systematically analyse how parliamentary discussion changes in response to different types of events in Australian history. We first compile a dataset of what was said in Australian state and federal parliaments from the mid-1800s through to 2017 based on available public records. We then use a structural text model to reduce the dimensionality of the text and to impose prior knowledge such as correlation between days and steady changes over time. Finally, we consider the effect of various events using Bayesian hierarchical Dirichlet regression. We find that: 1) elections are associated with topic changes only when the party in power changes; 2) a change in government can be associated with topic changes even if the party in power does not change; 3) economic events, such as financial crises, have less significant and persistent effects than other events such as terrorist attacks. Our findings have implications for how we think about the longer-term trajectory of government policymaking as the media cycle becomes increasingly focused on the short-term."
keywords: "text analysis, politics, Australia"
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt
toc: FALSE
# spacing: double
bibliography: ../bibliography.bib
biblio-style: apsr
endnote: no
---

@@

# Introduction
Government policy is partly driven by parliamentary discussion. Conversely, that same discussion can indicate a government's priorities. But major events--both expected, such as an election, and unexpected, such as a recession or terrorist attack--can affect the course of parliamentary discussion. For instance, think of how new governments often go to some trouble to be different to the one they replace. Or think of how events such as the 9/11 attacks altered government priorities.

In this paper we examine text records of what was said in Australian state and federal parliaments. Our earliest record is from 1851 and we consider the time period through to 2017. We use the Structural Topic Model (STM) of @RobertsStewartAiroldi2016 for dimensionality reduction and to impose structure such as correlation between days and changes over time. We then use Bayesian hierarchical Dirichlet regression to examine changes at various types of events, including: changes in government; changes in the political environment (as defined by polling or other results); changes in economic conditions; and other significant events (such as the 9/11 attacks or the Bali bombings).

We find: 1) elections are associated with topic changes only when the party in power changes; 2) a change in government can be associated with topic changes even if the party in power does not change; 3) economic events, such as financial crises, have less significant and persistent effects than other events such as terrorist attacks.

Our paper applies a topic model to a dataset of larger-scale parliamentary text records from multiple Australian parliaments. The Bayesian hierarchical Dirichlet regression allows us to construct a counterfactual. Our work fits into a growing literature that considers text as an input to more usual quantitative techniques, rather than requiring separate analysis. While using text as data has well-known shortcomings, it allows larger-scale analysis that would not be viable using less-automated approaches and so it can identify patterns that may otherwise be overlooked.

@@

Australian parliaments generally make their daily Hansard records available online as PDFs and these are considered the official release. Additionally, there is a more limited set of XML records available in some cases.[^XMLfootnote] We detail the sources of our Hansard PDFs in \@ref(hansardsources) and Appendix \@ref(examplehansardpage) provides an example of what a Hansard PDF looks like. There are roughly 54,331 **[UPDATE]** days worth of publicly available Hansard records across the state and federal parliaments (Table \ref{tab:yearsusedtable}). Our data cleaning process indicates concerns with a small number of PDFs and these are detailed in Appendix \@ref(knownhansardissues).

@@

## Additional information
Data from other sources are needed to complement the parliamentary text data. For instance, although the name, party, and division represented are contained in Hansard records they are inconsistent. Also non-Hansard information about the politicians informs our analysis. This includes: date of birth and date of death; sex; date of entry to parliament and exit from parliament; some aspects of their pre- and post-parliamentary career; some aspects of their parliamentary career, such as ministry appointments; and electoral information such as primary and two-party-preferred results.

@@

# Model
The main model that we use in this paper is the Structural Topic Model (STM) as implemented by the \texttt{stm} R package of @RobertsStewartAiroldiRPackage and a Bayesian discrete choice model. In a similar way to @MuellerRauh2018, we use topics to analyse the effect of various events. In Appendix \@ref(word2vec), we include an alternative approach that follows @Taddy2015 by using word2vec, which more closely uses words, rather than topics, as an input.

The basis of the STM is the Latent Dirichlet Allocation (LDA) model of @Blei2003latent. In this section we provide a brief overview of both the LDA model and the STM. We consider the outputs of the topic model as reduced dimension inputs that can be analysed within another model. In our case they are inputs to a discrete choice model, and so we then discuss that model.

@@

### Considering events
We are interested in the effect of political, economic, and other events on Australian parliamentary discussion. Political events are those related to a change of government or an election. Economic events are defined by substantial changes in various economic measures, such as the onset of the Great Depression or the Great Recession. Other events are those such as the 9/11 attacks, or the Bali Bombing. The full list of events that we consider is detailed in Appendix @ref(eventdetails).

There are many ways to statistically consider the effect of events. One way would be to ignore the time series aspect and just group the topic proportions by each event and then run multinomial tests for whether the groups are different. Another would be to consider some type of multinomial logistic regression. Options that would take advantage of the time series nature of our data include: autoregressive moving average (ARMA) models which can be used to test for structural breaks  **[REFERENCE?]**; or splines could be fit with specific knot placements and then cross validation used to compare the RSS with a more general splines model.

**[MONICA  - LOOK AT THESE OR OTHER OPTIONS AND DECIDE ON A WAY. NEEDS TO ALLOW FOR SHORT AND LONG RUN AND ALSO ALLOW COINCIDENCE OF EVENTS. WE ONLY CONSIDER 20 TOPICS BELOW, BUT THAT'S UNDERDONE AND THE METHODS NEED TO ALLOW FOR MORE.]**

We currently consider events in two ways. The first is graphically, grouping each of the daily topic proportions by the events and isolating the topics. This illustrates the differences between the topics. The second is by including incrementing variables in the prevalence metadata This groups each day by events and we can then conduct significance tests. Unfortunately both of these do not consider short-term effects.

**[Use tf-idf between them as illustrative instead? Use that other measure of difference?]**
<!-- We consider events in two ways. The first is in the change in the word usage before and after events and the second is in changes in the topics.  -->

@@

# Results
We are interested in considering the effect of various events on what is talked about in Australia's parliaments. Each of those parliaments are treated independently here. Future work could expand the model to better understand, and allow, for correlation between them.

@@

@@

## Selection of number of topics {#selecttopicnumber}
**[TBD]**

@@

governments %>% 
  kable(booktabs = T, caption = "Change in governments") %>%
  kable_styling(font_size = 12)

@@

elections %>% 
  filter(election == 1) %>% 
  select(year, electionDate, electionWinner) %>% 
  kable("latex", booktabs = T, caption = "Elections") %>%
  kable_styling(font_size = 12)

@@

keyEconomicDates %>% 
  kable(booktabs = T, caption = "Key economic events") %>%
  kable_styling(font_size = 12)

@@

keyOtherDates %>% 
  kable(booktabs = T, caption = "Key other events") %>%
  kable_styling(font_size = 12)

@@

>>>>

---
output:
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: false
    fig_caption: true
    number_sections: true
    latex_engine: pdflatex
    template: svm-latex-ms.tex
title: "The Effect of Events on Australian Parliamentary Discussion (1851--2017)"
thanks: "Thank you to John Tang, Zach Ward, Tim Hatton, Martine Mariotti, Tianyi Wang, Matt Jacob, Leslie Root, Jill Sheppard, Matthew Kerby, Chris Cochrane and Dan Simpson for their helpful suggestions; and to the UC Berkeley Demography Department for the use of their computing resources. Comments on the `r format(Sys.time(), '%d %B %Y')` version of this paper welcome at: rohan.alexander@anu.edu.au."
author:
- name: Monica Alexander
  affiliation: University of Toronto
- name: Rohan Alexander
  affiliation: Australian National University
abstract: "We systematically analyse how parliamentary discussion changes in response to different types of events in Australian history. We first create a dataset of what was said in Australian state and federal parliaments from the mid-1800s through to 2017 based on available public records. To reduce the dimensionality of this dataset we then use a topic model that allows correlation and a steady change over time. Finally, we consider the effect of various events using a Bayesian hierarchical Dirichlet model. We find that: 1) elections are associated with topic changes only when the party in power changes; 2) a change in government can be associated with topic changes even if the party in power does not change; 3) economic events, such as financial crises, have less significant and persistent effects than other events such as terrorist attacks. Our findings have implications for how we think about the longer-term trajectory of government policymaking as the media cycle becomes increasingly focused on short-term events."
keywords: "text analysis, politics, Australia"
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt
toc: FALSE
# spacing: double
bibliography: ../bibliography.bib
biblio-style: apsr
endnote: no
---

@@

# Introduction
Government policy is partly driven by parliamentary discussion. Conversely, that same discussion can indicate a government's priorities. But major events--both expected, such as an election, and unexpected, such as a recession or terrorist attack--can affect the course of parliamentary discussion. For instance, think of how a new government often goes to some trouble to appear different to the one they replace, or how events such as the 9/11 attacks altered government priorities.

In this paper we examine text records of what was said in Australian state and federal parliaments. Our earliest record is from 1851 and we consider the time period through to 2017. The dataset is detailed in Section \@ref(data). We use a topic model for dimensionality reduction and to allow for correlation between topics and steady changes over time. We then analyse the topics using a Bayesian hierarchical Dirichlet model to examine changes at various types of events. These events include: changes in government or elections; changes in economic conditions; and other significant events such as the 9/11 attacks. The models are detailed in Section \@ref(model).

Section \@ref(results) details our results, but broadly we find: 1) elections are associated with topic changes only when the party in power changes; 2) a change in government can be associated with topic changes even if the party in power does not change; 3) economic events, such as financial crises, have less significant and persistent effects than other events such as terrorist attacks. **[More here on the nature of the differences.]**

In this paper we apply a topic model to a dataset of larger-scale parliamentary text records from multiple Australian parliaments and then use a Bayesian hierarchical Dirichlet model to examine events. From a data perspective we contribute **[X]**. From a methods perspective, we contribute **[Y]**. Our model ican be used in many settings where the results of topic models need to be further analysed.

Our work is cross-disciplinary and in the traditions of political economy, economic history, political science and statistics. It contributes a growing modern quantitative social sciences literature that considers text as an input to more usual techniques, rather than requiring separate analysis. As the digitisation of historical sources continues and computational power becomes cheaper, we expect interest in this approach to only increase.

@@

Australian parliaments generally make their daily Hansard records available online as PDFs and these are considered the official release. Additionally, XML records are available in some cases.[^XMLfootnote] We detail the sources of our Hansard PDFs in \@ref(hansardsources) and Appendix \@ref(examplehansardpage) provides an example of what a Hansard PDF looks like. There are roughly 54,331 **[UPDATE]** days worth of publicly available Hansard records across the state and federal parliaments (Table \ref{tab:yearsusedtable}). Our data cleaning process indicates concerns with a small number of PDFs and these are detailed in Appendix \@ref(knownhansardissues).

@@

## Additional information
**[Not sure this is being used - might be better in the other paper.]**
Data from other sources are needed to complement the parliamentary text data. For instance, although the name, party, and division represented are contained in Hansard records they are inconsistent. Also non-Hansard information about the politicians informs our analysis. This includes: date of birth and date of death; sex; date of entry to parliament and exit from parliament; some aspects of their pre- and post-parliamentary career; some aspects of their parliamentary career, such as ministry appointments; and electoral information such as primary and two-party-preferred results.

@@

# Model
The models that we use in this paper are the Structural Topic Model (STM) as implemented by the \texttt{stm} R package of @RobertsStewartAiroldiRPackage, and Bayesian hierarchical Dirichlet analysis model. In a similar way to @MuellerRauh2018, we use topics as an input to another model, in our case to analyse the effect of various events. In Appendix \@ref(word2vec), we include an alternative approach that follows @Taddy2015 by using word2vec, which more closely uses words, rather than topics, as an input.

The basis of the STM is the Latent Dirichlet Allocation (LDA) model of @Blei2003latent. In this section we provide a brief overview of both the LDA model and the STM. We consider the outputs of the topic model as reduced dimension inputs that can be analysed within another model. We then discuss the Bayesian hierarchical Dirichlet analysis model that we use for this purpose.

@@

## Analysis model
We are interested in the effect of political, economic, and other events on Australian parliamentary discussion. Political events are those related to a change of government or an election. Economic events are defined by substantial changes in various economic measures, such as the onset of the Great Depression or floating the currency. Other events are those such as the 9/11 attacks. The full list of events that we consider are detailed in Appendix \@ref(eventdetails).

<!-- **[Actually, maybe this whole paragraph not needed?]** There are many ways to consider the effect of events. One is to summarise the topic proportions by each event and then run multinomial tests for whether the groups are different, although this does not take full advantage of the frequency of our data. Another is to consider a variant of multinomial logistic regression, but then events would be constrained to have an effect in certain ways. An autoregressive moving average (ARMA) model could be used to test for structural breaks, but **[something]**. Or splines could be fit with specific knot placements and then cross validation used to compare the RSS with a more general splines model, but **[something]**.  -->

One output of topic modelling is a probability distribution over the topics for each document. In our case the documents are each day's text. We analyse events by focusing on those proportions, $\gamma$, which we index by sitting period $s$, time $t$ and topic $p$, that is $\gamma_{stp}$. Note that time $t$ refers to the day index within a sitting period which is any collection of days separated from other days by at least a week. Our analysis model is:

\begin{align}
\gamma_{stp} &\sim \mbox{Dirichlet}(\mu_{stp}) (\#eq:distributionforgammas) \\ 
\log \mu_{stp} &= \alpha_{sp} + \delta_{stp}. (\#eq:meanforgammas) \\
\alpha_{s1} &= 0 \\
\alpha_{sp} &\sim \mbox{Normal}\left(\eta_{g[s]}, \sigma^2_g\right) \mbox{ for } p>1 \\
\delta_{s,t,p} &\sim \mbox{Normal}\left(\delta_{s,t-1,p}, \sigma^2_{\delta}\right) \\
\sigma^2_{\delta} &\sim \mbox{Uniform}(0, 40) 
\end{align}

Equation \@ref(eq:distributionforgammas) describes the proportion of each day within a sitting period given to a particular topic as a draw from a Dirichlet distribution parameterised by $\mu_{stp}$. The Dirichlet distribution is the multivariate extension of the Beta distribution, and the parameter controls where the 'weight' of the distribution lies. Equation \@ref(eq:meanforgammas) describes that distribution parameter, $\mu_{stp}$, as the sum of $\alpha_{sp}$ which is **X**, and $\delta_{stp}$, which is **Y**. These are themselves parameters. The primary aim of the modelling task is to find an appropriate and meaningful functional form for $\mu$. We consider $\alpha_{sp}$ as a draw from a Normal distribution with a mean that depends on **A** and a variance that depends on **B**. Similarly, $\delta_{stp}$ is modelled as a draw from a Normal distribution with a mean that depends on the preceding $\delta$ and some variance which is drawn from the Uniform distribution.

**[Does the variance parameter used in Equation 4 need to be specified too?]**

An illustration of the validity of our analysis model using simulated data is in Appendix \@ref(modeldetailsandsimulation).

<!-- We currently consider events in two ways. The first is graphically, grouping each of the daily topic proportions by the events and isolating the topics. This illustrates the differences between the topics. The second is by including incrementing variables in the prevalence metadata This groups each day by events and we can then conduct significance tests. Unfortunately both of these do not consider short-term effects. -->

<!-- **[Use tf-idf between them as illustrative instead? Use that other measure of difference?]** -->

<!-- We consider events in two ways. The first is in the change in the word usage before and after events and the second is in changes in the topics.  -->

@@

# Results
We are interested in considering the effect of various events on what is talked about in Australia's parliaments. Each of the parliaments and their houses are treated independently here. Future work could expand the model to better understand, and allow, for correlation between them.

@@

While using text as data has well-known shortcomings, it allows larger-scale analysis that would not be viable using less-automated approaches and so it can identify patterns that may otherwise be overlooked.

@@

## Selection of number of topics {#selecttopicnumber}

**[TBD]**

@@

governments %>% 
  kable(booktabs = T, caption = "Change in governments") %>%
  kable_styling(font_size = 10)

@@

elections %>% 
  filter(election == 1) %>% 
  select(year, electionDate, electionWinner) %>% 
  kable("latex", booktabs = T, caption = "Elections") %>%
  kable_styling(font_size = 10)

@@

keyEconomicDates %>% 
  kable(booktabs = T, caption = "Key economic events") %>%
  kable_styling(font_size = 10)

@@

keyOtherDates %>% 
  kable(booktabs = T, caption = "Key other events") %>%
  kable_styling(font_size = 10)

@@

# Analysis model {#modeldetailsandsimulation}

Here we illustrate the validity of our analysis model using simulated data.
